Weather and fuel as modulators of grassland fire behavior in the northern Great Plains

### 2022-08-24 paper/GrasslandFireBehaviorNGP.Rmd
start Env Man R1

---
title: "Weather and fuel as modulators of grassland fire behavior in the northern Great Plains"
author: 'Devan Allen McGranahan^A,C^, Megan E. Zopfi^B^, Kathryn A. Yurkonis^B^ ^A^ USDA Agricultural Research Service, Livestock and Range Research Laboratory, Miles City, Montana, USA ^B^ Department of Biology, University of North Dakota, Grand Forks, North Dakota, USA ^C^ Corresponding author. Email: Devan.McGranahan@usda.gov' 
output:
  word_document:
    reference_docx: word_template.docx
    keep_md: yes
csl: ijwf.csl
bibliography: FireBehaviorBib.bib
---

@@

Fuel and weather interact to affect wildland fire behavior, but little is known about associations between these variables in the northern Great Plains of North America. 
Few studies consider rate of spread or statistically test the influence of fire environmental variables.
We measured fuel load and moisture ahead of prescribed fires in North Dakota, USA, and used a multi-channel thermocouple array to measure rate of spread, soil surface temperature, and flame temperature, which we compared with fire weather data.
Aboveground flame temperatures averaged 225$^\circ$C during spring burns in and 250$^\circ$C during fall burns.
Surface temperatures averaged just above 100$^\circ$C, although 50% of observations were 60$^\circ$C or less.
Fires spread at an average of 2.5 m min$^{\text{-}1}$. 
Regression analysis indicated that wind speed drove faster rates of spread while higher fuel loads and lower overall fuelbed moisture produced higher flame temperatures.
None of our measurements explained variability in soil surface temperature, likely because ground-level heating remained low.
We highlight differential responses among fire behavior metrics to different components of the wildland fire environment.
Results help wildland fire managers match burn conditions to desired outcomes, and fire ecologists match measurements to specific ecological responses.
We also discuss remaining knowledge gaps in how management and ignition patterns interact to create variability in fire behavior.

@@

More than simply the result of combustion of vegetation, fire behavior in wildland environments is multi-faceted, with different components producing different effects on the surrounding environment and organisms within. 
Most wildland fire scientists describe fire behavior in terms of *rate of spread*—how quickly a flame front moves through a fuelbed—and *intensity*—a suite of measurements of how much energy is released by combustion, often expressed as a rate of energy release over time [@mcgranahan2021a].

Wildland fire behavior is controlled by interactions between several abiotic and biotic factors, and understanding them is critical to safe and effective wildland fire management [@benson2009]. 
*Abiotic factors* include those determined by the physical environment, such as wind speed and atmospheric moisture content. 
Wind speed has long been recognized as a primary driver of fire behavior, especially in well-cured grassland fuels [@whittaker1961; @cheney1995; @kidnie2015].
Two measures of atmospheric moisture content—relative humidity and vapor pressure deficit—are also associated with fire growth [@evett2008; @reid2010; @sedano2014].

*Biotic factors* relate principally to the amount and nature of plant biomass available for combustion. 
Overall, energy release rates increase as more fuel is available to burn; the structure and arrangement of vegetation is also important. 
Greater fuel load attributable to longer time-since-fire increases fire temperature, and spatial variability in fuel load and patchy distribution of fine fuels in turn drive variability in fire behavior [@patten1984; @gibson1990; @gomes2020].
Furthermore, fine-leaved grasses burn more completely and hotter than an equal mass of forbs [@wragg2018].
Finally, fuel moisture content is an especially important driver outside of the highly-cured context of wildfire seasons [@sparling1966; @kidnie2015]. 
Together, variability in flammability traits and curing rates among species that comprise grassland fuelbeds contributes to variability in fire behavior [@cruz2015; @kidnie2015; @mcgranahan2016; @cardoso2018].

@@

Our objectives were to (1) describe the range of variability in three measures of fire behavior—rate of spread, soil surface temperature, and flame temperature 15 cm above soil surface—during prescribed burns in typical fuelbeds of the northern US Great Plains, and (2) explain variability in fire behavior in terms of abiotic and biotic conditions. 
Our analysis emphasizes the differential effects of environmental variables among the three responses, and the multidimensional relationship among these different measures of grassland fire behavior during prescribed burns conducted at scales consistent with land management in the region.

@@

We measured each fire with a set of 9, 1 m equilateral triangle plots arranged in a nested fashion such that three 10 m triangles, each containing 3, 1m plots, were placed 100 m apart to form a total plot area of 0.433 ha with 27 sample points positioned at the centroid of each burn unit (A detailed schematic of this layout is presented in Supplemental Information Figure 2). 
This fractal design is modified from the Sierpinksi triangle described by @dorrough2007 and applied to measuring wildland fire spread by @mcgranahan2021b.
Although the logarithmically-scaled nested design was intended for geospatial analysis of point-level data, for our analyses here we calculate averages from the finest (1 m) scale, consistent with the method of locating multiple microplots within larger burned areas to characterize spatial variability within fires [@fernandes2000].

@@

**Fire behavior data** were recorded as temperature ($^\circ$C) associated with the advancing flame front at each of the 27 points arranged in 9, 1 m triangles at the center of each burn unit.
Data were recorded with the open-source FeatherFlame thermocouple datalogger system [@mcgranahan2021b], logging at 1.5 Hz.
Briefly, the FeatherFlame system reads overbraided, ceramic fiber-insulated, 20-gauge K-type thermocouples (Omega, Norwalk, CT) connected to an Arduino-based datalogger assembled from Adafruit Feather breakout boards (M0 Adalogger, datalogging shield, and OLED display; Adafruit Industries, LLC, New York City, NY) and housed inside water-resistant Pelican cases (Pelican Products, Inc, Torrance, California). 
The low cost of open-source systems make multiple units more affordable than proprietary data loggers with no sacrifice in data quality [@mcgranahan2021].

At each 1-m triangle—the individual observational unit in the nested plot design—we used four thermocouples connected to a single FeatherFlame datalogger.
Three thermocouples measured flame temperature 15 cm above the soil surface at each 1 m vertex while a fourth thermocouple recorded soil surface temperature at a representative point within the 1 m array.
Beads of the thermocouple probes extended at least 3 cm from the supporting apparatus, to which the insulated lead was attached with wire. 
The soil thermocouple was placed on mineral ground, perpendicular to the soil surface, below plant litter.

For each fire event, we determined the time and value of the maximum temperature ($^\circ$C) as the flame front encountered each thermocouple. 
We calculated the maximum flame temperature ($^\circ$C) 15 cm above soil surface for each plot as the mean of the thermocouple readings from its vertices. 
We calculated the rate of spread (m min$^{\text{-}1}$) of the flame front as it passed through each plot using the maximum temperature timestamps following equations from @simard1984, which are presented in full in Supplemental Information.

@@

Prior to analysis, we ensured statistical power across 167 observational units by using multiple imputation to interpolate missing datapoints, as missing field data occurred for three rate of spread samples (2% of total), 29 fuel load values (17%), and 46 soil surface temperature values (27%). 
We used the multiple imputation method in the *mice* package [@vanbuuren2011] in the R statistical environment [@rct2020] to fill in these missing values. 
The procedure simulated 50 datasets with different, but reasonable, values for the missing data based on patterns in the existing data. 
We then scaled all variables to a common range within each imputed dataset, performed regression analysis on each *mice*-generated dataset, and report composite statistical results pooled from the results of the 50 individual regression models.

@@

**Regression analysis.---**We assessed weather and fuel effects on three fire behavior response variables: Maximum flame temperature ($^\circ$C) 15 cm above the soil surface (mean of three thermocouples), maximum soil surface temperature ($^\circ$C; single thermocouple), and rate of spread (m min$^{\text{-}1}$) through each 1 m equilateral triangle plot. 
Because all three response variables were best modeled with a gamma distribution, we fit generalized linear mixed-effect regression models for each response with the 'glmer' function from the *lme4* package in R [@bates2015]. 
Fixed effects consisted of weather and fuel variables, as described above. 
The random-effect term was constructed to account for random variance among locations, spatial non-independence within locations and nested variance within sample plots, and the effect of repeated measurements within each location. 
Due to concerns about collinearity between relative humidity and vapor pressure deficit because they are derived from the same variables, vapor pressure deficit was excluded from regression analysis for all three response variables.

@@

Mean temperatures recorded in this study are consistent with other reports from northern rangelands. 
In central Alberta grassland, @bailey1980 observed that surface temperatures varied between 110$^\circ$C and 165$^\circ$C for backfires and headfires, respectively, and headfires averaged 200$^\circ$C 15 cm above the ground; temperatures generally tracked with fuel load. 
Surface fires through jack pine barrens in Ontario had a similar range as ours: 140-545$^\circ$C [@smith1966]. 
In our study, average 15-cm temperatures were 225$^\circ$C during spring burns in central North Dakota and 250$^\circ$C during fall burns in southwestern North Dakota; surface temperatures at both locations generally averaged just above 100$^\circ$C (Fig. 1).

Discrepancies between our data and others reported in the region are consistent with what would be expected when differences in the fire environment are considered.
For example, @ohrtman2015 reported a wide range of maximum temperatures at the soil surface---150-500$^\circ$C---that was generally explained by variability in annual productivity and clipping frequency. 
Our fires were also cooler than those reported by @archibold2003 in Saskatchewan: using the mid-point of observations made at 10 cm and 20 cm as a comparison, spring fires reached 314$^\circ$C and fall fires reached 298$^\circ$C. 
But @archibold2003 also reported substantially lower fuel moisture in each season and they had approximately three times the fuel load, likely due to an absence of grazing management on the remnant prairie. 
A previous study reported similar results---temperatures approaching 500$^\circ$C when fuel loads averaged 2.8-4.5 t ha$^{\text{-}1}$ [@archibold1998]. 
With greater variability in fuel load and fuel moisture, we might also expect to see these factors have greater influence on aboveground flame temperatures. 
For example, in Colorado, @augustine2014 observed a strong linear relationship between fuel load and  temperatures ranging from 60-200$^\circ$C, but their fuel load also ranged from 0.2 to 1.2 t ha$^{\text{-}1}$.

Fires at both of our locations spread much more slowly than most reports from other grasslands, due primarily to high fuelbed moisture content and little opportunity for mitigation by wind or dry air (Fig. 1). 
Perhaps the most variability in fire spread was reported by @sneeuwjagt1977 from prescribed grass fires in California and Washington, where rates of spread ranged from 0.2-61 m min$^{\text{-}1}$.
From grassland fires in South Africa and Kansas, USA, @trollope2002 reported average spread rates of 24 and 32 m min$^{\text{-}1}$, respectively, for head fires and 0.12 and 0.14 m min$^{\text{-}1}$, respectively, for back fires. 
In a tallgrass prairie in Texas, @clements2019 recorded fire spreading between 72-150 m min$^{\text{-}1}$ with the wind and 48 m min$^{\text{-}1}$ for flanking fires.
Likewise, fires through cured grass fuelbeds in Australia spread much more rapidly than we observed---up to 18-180 m min$^{\text{-}1}$ [@cheney1993; @cheney1995; @cruz2015]. 
Through partially-cured stands, though, @cruz2015 found spread rates dropped to 3-44 m min$^{\text{-}1}$, approaching our location averages of 2.1 and 3.1 m min$^{\text{-}1}$. 
Consistent with our finding that only wind speed had a statistically-significant effect on increasing rate of spread (Fig. 3), @cheney1993 found that wind was by far the most important variable to spread rate.

@@

This study is novel in that it examines the fire environment at a spatial scale consistent with land management in the region using realistic ignition scenarios. 
To our knowledge, no other study in the northern Great Plains has reported the behavior of fires larger than experimental plots. 
Integrating research into management almost invariably requires trade-offs; two already discussed above include (1) measuring only the overall moisture content of the entire fuelbed, being precluded from parsing fuel into live, dead, and litter components, and (2) measuring two-dimensional rate of spread of fire fronts within the burn unit without being able to associate them with wind direction. 
But these are the respective conditions under which prescribed fire managers in the region decide whether to burn, and ensure fire spread objectives are met. 
Research conducted at the spatial scales at which management occurs helps managers trust the transfer of knowledge from studies to working landscapes [@sayre2005; @cacciapaglia2012].
For example, all of our fire behavior measurements were made more than 50 m from the initial fire line, the distance identified in simulations and used in wildland fire science to allow flame fronts to achieve a quasi steady state in spread rate [@fernandes2000; @sutherland2020], which is obviously precluded in studies that employ small plots.

>>>>

---
title: Weather and fuel as modulators of grassland fire behavior in the northern Great
  Plains
author: 'Devan Allen McGranahan^A,C^, Megan E. Zopfi^B^, Kathryn A. Yurkonis^B^ ^A^
  USDA Agricultural Research Service, Livestock and Range Research Laboratory, Miles
  City, Montana, USA ^B^ Department of Biology, University of North Dakota, Grand
  Forks, North Dakota, USA ^C^ Corresponding author. Email: Devan.McGranahan@usda.gov'
output:
  pdf_document: 
     citation_package: natbib
     keep_tex: yes
  word_document:
    reference_docx: word_template.docx
    keep_md: yes
csl: ijwf.csl
bibliography: FireBehaviorBib.bib
---

@@

**Background:** 
Fuel and weather interact to affect wildland fire behavior, but little is known about associations between these variables in the northern Great Plains of North America. 
Few studies consider rate of spread or statistically test the influence of fuel and weather.

**Methods:** 
We measured overall fuel load and moisture ahead of prescribed fires in North Dakota, USA, and used a thermocouple array to measure rate of spread, soil surface temperature, and aboveground flame temperature, to compare with fire weather data.

**Key results:** 
Flame temperatures averaged 225$^\circ$C during spring burns and 250$^\circ$C during fall burns, and were generally higher with greater fuel loads and lower overall fuelbed moisture. 
Surface temperatures averaged ~100$^\circ$C, although 50% of observations were ≤60$^\circ$C. 
Fires spread at an average of 2.5 m min$^{\text{-}1}$, increasing with wind speed.

**Conclusions:** 
Prescribed fire in northern Great Plains working rangeland spreads slowly and effects low soil surface temperatures, often limited by high fuelbed moisture. 
Fire behavior measurements respond differently to variability in fuel and weather. Belowground heating is likely minimal.

**Implications:** 
Ecologists, consider which fire behavior measurements best relate to fire effects. 
Managers, consider weather and ignition pattern mitigations when fuels constrain desired fire behavior to ensure effective burns.

@@

More than simply the result of combustion of vegetation, wildland fire behavior is multi-faceted, with different components producing different effects on the surrounding environment and organisms within. 
Most wildland fire scientists describe fire behavior in terms of *rate of spread*—how quickly a flame front moves through a fuelbed—and *intensity*—a suite of measurements of how much energy is released by combustion, often expressed as a rate of energy release over time [@mcgranahan2021a].

Wildland fire behavior is controlled by interactions among several abiotic and biotic factors, and understanding them is critical to safe and effective wildland fire management [@benson2009]. 
*Abiotic factors* include those determined by the physical environment, such as wind speed and atmospheric moisture content. 
Wind speed has long been recognized as a primary driver of fire behavior, especially in well-cured grassland fuels [@whittaker1961; @cheney1995; @kidnie2015].
Two measures of atmospheric moisture content—relative humidity and vapor pressure deficit—are also associated with fire growth [@evett2008; @reid2010; @sedano2014].

*Biotic factors* relate principally to the amount and nature of plant biomass available for combustion. 
Overall, energy release rates increase as more fuel is available to burn. 
The structure and arrangement of vegetation is also important. 
Greater fuel load attributable to longer time-since-fire increases fire temperature, and spatial variability in fuel load and patchy distribution of fine fuels in turn drive variability in fire behavior [@patten1984; @gibson1990; @gomes2020].
Furthermore, fine-leaved grasses burn more completely and hotter than an equal mass of forbs [@wragg2018].
Finally, fuel moisture content is an especially important driver outside of the highly-cured context of wildfire seasons [@sparling1966; @kidnie2015]. 
Together, variability in flammability traits and curing rates among species that comprise grassland fuelbeds contributes to variability in fire behavior [@cruz2015; @kidnie2015; @mcgranahan2016; @cardoso2018].

@@

Our objectives were to (1) describe the range of variability in three measures of fire behavior—rate of spread, soil surface temperature, and flame temperature 15 cm above soil surface—during prescribed burns in typical fuelbeds of the northern US Great Plains, and (2) explain variability in fire behavior in terms of abiotic and biotic factors.  
Our analysis emphasizes the differential effects of environmental variables among the three measures of fire behavior, and the multidimensional relationship among these responses during prescribed burns conducted at scales consistent with land management in the region.

@@

We measured each fire with a set of 9, 1-m equilateral triangle plots arranged in a nested fashion such that three 10 m triangles, each containing 3, 1-m plots, were placed 100 m apart to form a total plot area of 0.433 ha with 27 sample points positioned at the centroid of each burn unit (a schematic of this layout is presented in Supplemental Information Figure 2). 
This fractal design is modified from the Sierpinksi triangle described by @dorrough2007 and applied to measuring wildland fire spread by @mcgranahan2021b.
Although the logarithmically-scaled nested design was intended for geospatial analysis of point-level data, for our analyses here we calculate averages from the finest (1 m microplot) scale, consistent with the method of locating multiple microplots within larger burned areas to characterize spatial variability within fires [@fernandes2000].

@@

**Fire behavior data** were recorded as temperature ($^\circ$C) associated with the advancing flame front at each of the 27 points arranged in 9, 1-m triangular microplots at the center of each burn unit.
Data were recorded with the open-source FeatherFlame thermocouple datalogger system [@mcgranahan2021b], logging at 1.5 Hz.
The FeatherFlame system reads overbraided, ceramic fiber-insulated, 20-gauge K-type thermocouples (Omega, Norwalk, CT) connected to an Arduino-based datalogger assembled from Adafruit Feather breakout boards (M0 Adalogger, datalogging shield, and OLED display; Adafruit Industries, LLC, New York City, NY) and housed inside water-resistant Pelican cases (Pelican Products, Inc, Torrance, California).
The low cost of open-source systems make multiple units more affordable than proprietary data loggers with no sacrifice in data quality [@mcgranahan2021]. 
More details on the datalogger system are available in Supplemental Information.

At each 1 m triangular microplot—the observational unit in the nested plot design—we used four thermocouples connected to a single FeatherFlame datalogger.
Three thermocouples measured flame temperature 15 cm above the soil surface at each 1 m vertex while a fourth thermocouple recorded soil surface temperature at a representative point within the 1 m array.
Beads of the thermocouple probes extended at least 3 cm from the supporting apparatus, to which the insulated lead was attached with wire. 
The soil thermocouple was placed on mineral ground, perpendicular to the soil surface, below plant litter.

For each fire event, we determined the time and value of the maximum temperature ($^\circ$C) as the flame front encountered each thermocouple. 
We calculated the maximum flame temperature ($^\circ$C) 15 cm above soil surface for each microplot as the mean of the thermocouple readings from its 3 vertices. 
We calculated the rate of spread (m min$^{\text{-}1}$) of the flame front as it passed through each microplot using the maximum temperature timestamps as arrival times following equations from @simard1984, which are presented in full in Supplemental Information.

@@

Prior to analysis, we ensured statistical power across 167 observational units by using multiple imputation to interpolate missing datapoints, as missing field data occurred for three rate of spread samples (2% of total), 29 fuel load values (17%), and 46 soil surface temperature values (27%) due to logistical and time constraints during the operational burn periods. 
We used the multiple imputation method in the *mice* package [@vanbuuren2011] in the R statistical environment [@rct2020] to fill in these missing values. 
The procedure simulated 50 datasets with different, but reasonable, values for the missing data based on patterns in the existing data. 
We then scaled all variables to a common range within each imputed dataset, performed regression analysis on each *mice*-generated dataset, and report composite statistical results pooled from the results of the 50 individual regression models.

@@

**Regression analysis.---**We assessed weather and fuel effects on three fire behavior response variables: Maximum flame temperature ($^\circ$C) 15 cm above the soil surface (mean of three thermocouples), maximum soil surface temperature ($^\circ$C; single thermocouple), and rate of spread (m min$^{\text{-}1}$) through each 1 m trianglular microplot. 
Because all three response variables were best modeled with a gamma distribution, we fit generalized linear mixed-effect regression models for each response with the 'glmer' function from the *lme4* package in R [@bates2015]. 
Fixed effects consisted of weather and fuel variables, as described above. 
The random-effect term was constructed to account for random variance among locations, spatial non-independence within locations and nested variance within sample plots, and the effect of repeated measurements within each location. 
Due to concerns about collinearity between relative humidity and vapor pressure deficit because they are derived from the same variables, vapor pressure deficit was excluded from regression analysis for all three response variables.

@@

Mean temperatures recorded in this study are consistent with other reports from northern rangelands. 
In central Alberta grassland, @bailey1980 observed that surface temperatures varied between 110$^\circ$C and 165$^\circ$C for backfires and headfires, respectively, and headfires averaged 200$^\circ$C 15 cm above the ground; temperatures generally tracked with fuel load. 
Surface fires through jack pine barrens in Ontario had a similar range as ours [140-545$^\circ$C, @smith1966]. 
In our study, mean 15-cm temperatures were 225$^\circ$C during spring burns in central North Dakota and 250$^\circ$C during fall burns in southwestern North Dakota; surface temperatures at both locations generally averaged just above 100$^\circ$C (Fig. 1).

Discrepancies between our data and others from the region are consistent with what would be expected when differences in the fire environment are considered.
For example, @ohrtman2015 reported a wide range of maximum temperatures at the soil surface---150-500$^\circ$C---that was generally explained by variability in annual productivity and clipping frequency, which altered fuel load. 
Our fires were also cooler than those reported by @archibold2003 in Saskatchewan: using the mid-point of observations made at 10 cm and 20 cm as a comparison to our 15-cm values, spring fires reached 314$^\circ$C and fall fires reached 298$^\circ$C. 
But @archibold2003 also reported substantially lower fuel moisture in each season and they had approximately three times the fuel load, likely due to an absence of grazing on the remnant prairie. 
A previous study reported similar results---temperatures approaching 500$^\circ$C when fuel loads averaged 2.8-4.5 t ha$^{\text{-}1}$ [@archibold1998]. 
With greater variability in fuel load and fuel moisture, we might also expect to see these factors have greater influence on aboveground flame temperatures. 
For example, in Colorado, @augustine2014 observed a strong linear relationship between fuel load and  temperatures ranging from 60-200$^\circ$C, but their fuel load also ranged from 0.2 to 1.2 t ha$^{\text{-}1}$.

Fires at both of our locations spread much more slowly than most reports from other grasslands, due primarily to high fuelbed moisture content and little opportunity for mitigation by wind or lower atmospheric moisture (Fig. 1). 
Perhaps the most variability in fire spread was reported by @sneeuwjagt1977 from prescribed grass fires in California and Washington, where rates of spread ranged from 0.2-61 m min$^{\text{-}1}$.
From grassland fires in South Africa and Kansas, USA, @trollope2002 reported average spread rates of 24 and 32 m min$^{\text{-}1}$, respectively, for head fires and 0.12 and 0.14 m min$^{\text{-}1}$, respectively, for back fires. 
In a tallgrass prairie in Texas, @clements2019 recorded fire spreading between 72-150 m min$^{\text{-}1}$ with the wind and 48 m min$^{\text{-}1}$ for flanking fires.
Likewise, fires through cured grass fuelbeds in Australia spread much more rapidly than we observed---up to 18-180 m min$^{\text{-}1}$ [@cheney1993; @cheney1995; @cruz2015]. 
Through partially-cured stands, though, @cruz2015 found spread rates dropped to 3-44 m min$^{\text{-}1}$, approaching our location averages of 2.1 and 3.1 m min$^{\text{-}1}$. 
Consistent with our finding that only wind speed had a statistically-significant effect on increasing rate of spread (Fig. 3), @cheney1993 found that wind was by far the most important variable to spread rate.

@@

This study is novel in that it examines the fire environment at a spatial scale consistent with land management in the region using realistic ignition scenarios. 
To our knowledge, no other study in the northern Great Plains has reported the behavior of fires larger than experimental plots. 
Integrating research into management almost invariably requires trade-offs; two already discussed above include (1) measuring only the overall moisture content of the entire fuelbed, being precluded from parsing fuel into live, dead, and litter components, and (2) measuring two-dimensional rate of spread of fire fronts within the burn unit without being able to associate them with wind direction. 
But these are the respective conditions under which prescribed fire managers in the region decide whether to burn, and ensure fire spread objectives are met. 
Research conducted at the spatial scales at which management occurs helps managers trust the transfer of knowledge from studies to working landscapes [@sayre2005; @cacciapaglia2012].
For example, all of our fire behavior measurements were made more than 50 m from the initial fire line, the distance identified in simulations and used in wildland fire science to allow flame fronts to achieve a quasi-steady state in spread rate [@fernandes2000; @sutherland2020], which is obviously precluded in studies that employ small plots.

### 2022-08-24 paper/SupplementaryInfo.Rmd
start Env Man R1

---
title: Weather and fuel as modulators of grassland fire behavior in the northern Great Plains
subtitle: Supplementary information
date: '`r Sys.Date()`'
author: DA McGranahan, ME Zopfi, KA Yurkonis
header-includes:
  - \usepackage{amsmath}
  - \renewcommand{\familydefault}{\sfdefault}
  - \usepackage{hyperref}
  - \usepackage{booktabs}
output:
  pdf_document:
    highlight: tango
---

>>>>

---
title: Weather and fuel as modulators of grassland fire behavior in the northern Great Plains
subtitle: Supplementary information
date: '`r Sys.Date()`'
author: DA McGranahan, ME Zopfi, KA Yurkonis
header-includes:
  - \usepackage{amsmath}
  - \renewcommand{\familydefault}{\sfdefault}
  - \usepackage{hyperref}
  - \usepackage{booktabs}
output:
  pdf_document:
    highlight: tango
    keep_tex: yes
---
